---
title: "SCCD Multistep modelling"
output: html_document
date: "2026-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(bayestestR)
library(dplyr)
library(tidyr)
library(patchwork)
```

```{r}
source("utils.R")
source("utils_MCMC.R")
source("TDCM_MCMC_empirical.R")
# load("TDCM_empirical_three_level_response.RData")
# data3L = data
```

```{r}
load("TDCM_empirical_data.RData")
data2L = data
```

```{r}
Q_matrix= data2L$Q_matrix
Y = data2L$Y
group = data2L$group
X_group = data2L$X_group
X_ind = data2L$X_ind
J = dim(Q_matrix)[1] # the number of questions: 25
K = dim(Q_matrix)[2] # the number of skills: 5
N = dim(Y)[1] # the number of respondents: 987
indT = dim(Y)[2] # the number of time points: 2
G = max(group) # the number of groups: 52
p = dim(X_group[[1]][[1]])[2]  # number of group covariates
m = dim(X_ind[[1]][[2]])[2]  # number of individual covariates (including intercepts)

## remove individual covariates
for (k in seq(K)){
  X_ind[[k]][[2]] <- X_ind[[k]][[2]][,1, drop = FALSE] # keep only the intercept
}

## remove group covariates
for (k in seq(K)){
  for (i in seq(2)){
     X_group[[k]][[i]] <- X_group[[k]][[i]][,1, drop = FALSE] # keep only the intercept
  }
}

empirical_data3L = list('Y' = Y, 
                      'X_group' = X_group, 
                      'X_ind' = X_ind, 
                      'group' = group, 
                      'Q' = Q_matrix, 
                      'K' = K,
                      'indT' = indT
                      )


MCMC_result_empirical_1_2 = MCMC_Ind_fit(empirical_data2L, M = 1000, priors=list(beta_prior = 2.5, sigma_omega = 5, v_prior = 2), beta_interact = FALSE, two_way_only = TRUE)

saveRDS(MCMC_result_empirical_1_2, file = "MCMC_result_empirical_1_2.rds")
```

```{r}
MCMC_result_empirical_1_2 = readRDS("MCMC_result_empirical_1_2.rds")
MCMC_result_empirical_1 = readRDS("MCMC_result_empirical_1.rds")
```


```{r}
beta_plot = function(MCMC_result_empirical){
  
  beta_estimate = MCMC_result_empirical$E_beta

  intercept = vector(length = J)
  main_effect = vector(length = J)
  for (j in seq(J)){
    intercept[j] = beta_estimate[[j]][1]
    main_effect[j] = beta_estimate[[j]][2]
  }
  
  
  # Combine into long format for ggplot
  J <- length(intercept)  # Define J explicitly
  beta_estimate <- data.frame(
    index = rep(seq_len(J), 2),
    value = c(intercept, main_effect),
    type = rep(c("Intercept", "Main Effect"), each = J)
  )
  
  # Plot
  p1 <- ggplot(data = beta_estimate, aes(x = index, y = value, fill = type, shape = type)) +
    geom_point(size = 2.5, color = "white") +
    xlab("Parameter Index (beta)") +
    ylab("Estimate") +
    labs(title = " Estimated value of Beta") +
    scale_fill_manual(values = c("Intercept" = "palevioletred2", "Main Effect" = "lightseagreen")) +
    scale_shape_manual(values = c("Intercept" = 21, "Main Effect" = 24)) +
    theme_minimal()
  p1
}
```

```{r}
J = 25
beta_plot(MCMC_result_empirical_1)
```
```{r}
load("TDCM_empirical_data.RData")

Q_matrix= data$Q_matrix
Y = data$Y
group = data$group
X_group = data$X_group
X_ind = data$X_ind
J = dim(Q_matrix)[1] # the number of questions: 25
K = dim(Q_matrix)[2] # the number of skills: 5
N = dim(Y)[1] # the number of respondents: 987
indT = dim(Y)[2] # the number of time points: 2
G = max(group) # the number of groups: 52
p = dim(X_group[[1]][[1]])[2]  # number of group covariates
m = dim(X_ind[[1]][[2]])[2]  # number of individual covariates (including intercepts)


## remove group covariates
for (k in seq(K)){
  for (i in seq(2)){
     X_group[[k]][[i]] <- X_group[[k]][[i]][,1, drop = FALSE] # keep only the intercept
  }
}

empirical_data = list('Y' = Y, 
                      'X_group' = X_group, 
                      'X_ind' = X_ind, 
                      'group' = group, 
                      'Q' = Q_matrix, 
                      'K' = K,
                      'indT' = indT
                      )

# MCMC_result_empirical = MCMC_Ind_fit(empirical_data, M = 1000, priors=list(beta_prior = 1, sigma_omega = 2.5, v_prior = 1), beta_interact = FALSE, two_way_only = TRUE)

MCMC_result_empirical_3 = MCMC_Ind_fit(empirical_data, M = 1000, priors=list(beta_prior = 2.5, sigma_omega = 5, v_prior = 2), beta_interact = FALSE, two_way_only = TRUE)

saveRDS(MCMC_result_empirical_3, file = "MCMC_result_empirical_3.rds")

```

```{r}
MCMC_result_empirical_3 = readRDS("MCMC_result_empirical_3.rds")
```


```{r}
significant_gamma = matrix(nrow = 0, ncol = 16)
for(k in 1:K) {
  for(z in 1:2) {
    for(s in 1:52) {
      # sig_ind = (abs(colMeans(MCMC_result_empirical_3$E_gamma_trace[[k]][[2]][,z,s,-1])) - 1.28 * apply(MCMC_result_empirical_3$E_gamma_trace[[k]][[2]][,z,s,-1], 2, sd)) > 0
      sig_ind = apply(MCMC_result_empirical_3$E_gamma_trace[[k]][[2]][,z,s,-1], 2, function(x){prod(hdi(x, 0.8))}) > 0
      significant_gamma = rbind(significant_gamma, c(k, z, sig_ind))
    }
  }
}
significant_gamma = as.data.frame(significant_gamma)
colnames(significant_gamma) = c("attribute", "forward", "Q2A", "Q2B", "Q17B", "GENDER", "LUNCH_STATUS", "ELL_STATUS", "SPED_STATUS", "P2", "P3A", "P3C", "P6", "P7F", "FAST_ATTEND_RATE", "FASTWORKS_ATTEND_TOTAL")
```

```{r}
significant_gamma = significant_gamma %>%
  group_by(attribute, forward) %>%
  summarise(across(everything(), sum))
significant_gamma
```
```{r}
s1t2_f_SEPD = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[1]][[2]][,1,,8]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s1t2_f_SEPD$School = as.factor(substr(s1t2_f_SEPD$School, 2, 4))

plot = ggplot(s1t2_f_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "turquoise") +
  theme_bw() +
  labs(title = "Effect of special education status on acquiring attribute 1") +
  scale_x_discrete(labels = NULL)
ggsave("plots/s1t2_f_SPED.pdf", plot, width = 8, height = 4)
plot
```



```{r}
s2t2_f_SEPD = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[2]][[2]][,1,,8]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s2t2_f_SEPD$School = as.factor(substr(s2t2_f_SEPD$School, 2, 4))

plot = ggplot(s2t2_f_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "turquoise3") +
  theme_bw() +
  labs(title = "Effect of special education status on acquiring attribute 2") +
  scale_x_discrete(labels = NULL)
ggsave("plots/s2t2_f_SPED.pdf", plot, width = 8, height = 4)
plot
```
```{r}
s3t2_f_SEPD = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[3]][[2]][,1,,8]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s3t2_f_SEPD$School = as.factor(substr(s3t2_f_SEPD$School, 2, 4))

plot = ggplot(s3t2_f_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "turquoise4") +
  theme_bw() +
  labs(title = "Effect of special education status on acquiring attribute 2") +
  scale_x_discrete(labels = NULL)
ggsave("plots/s3t2_f_SPED.pdf", plot, width = 8, height = 4)
plot
```



```{r}
s4t2_f_SEPD = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[4]][[2]][,1,,8]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s4t2_f_SEPD$School = as.factor(substr(s4t2_f_SEPD$School, 2, 4))

plot = ggplot(s4t2_f_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "seagreen") +
  theme_bw() +
  labs(title = "Effect of special education status on acquiring attribute 2") +
  scale_x_discrete(labels = NULL)
ggsave("plots/s4t2_f_SPED.pdf", plot, width = 8, height = 4)
plot
```
```{r}
s4t2_b_SEPD = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[4]][[2]][,2,,8]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s4t2_b_SEPD$School = as.factor(substr(s4t2_b_SEPD$School, 2, 4))

plot = ggplot(s4t2_b_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "seagreen") +
  theme_bw() +
  labs(title = "Effect of special education status on maintaining attribute 2") +
  scale_x_discrete(labels = NULL)
ggsave("plots/s4t2_b_SPED.pdf", plot, width = 8, height = 4)
plot
```


```{r}
s3t2_b_P7F = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[5]][[2]][,1,,4]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")

ggplot(s3t2_b_P7F, aes(x = School, y = Coefficient, fill = School)) +
  geom_violin(trim = FALSE) +               # Create the violin shapes
  geom_boxplot(width = 0.1, aes(fill = School), outlier.shape = NA) + # Optional: add a tiny boxplot inside
  theme_minimal() +                         # Clean background
  labs(title = "Hierarchical effect of group level covariates on \nindividual level impact of FASTWORK attendance for \nskill 2 forward transition") +
  scale_fill_brewer(palette = "Pastel1")
```
```{r}
s1t2_b_LUNCH = as.data.frame(MCMC_result_empirical_3$E_gamma_trace[[1]][[2]][,2,,6]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s1t2_b_LUNCH$School = as.factor(substr(s1t2_b_LUNCH$School, 2, 4))

plot = ggplot(s1t2_b_LUNCH, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "seagreen") +
  theme_bw() +
  labs(title = "Effect of free or reduced price lunch status on maintaining attribute 1") +
  scale_x_discrete(labels = NULL)
plot
```




```{r}
MCMC_result_empirical_2 = readRDS("MCMC_result_empirical_2.rds")
```

```{r}
significant_omega = matrix(nrow = 0, ncol = 6)
for(k in 1:K) {
  for(z in 1:2) {
    # sig_ind = (abs(colMeans(MCMC_result_empirical_2$E_omega_trace[[k]][[2]][,z,-1])) - 1.28 * apply(MCMC_result_empirical_2$E_omega_trace[[k]][[2]][,z,-1], 2, sd)) > 0
    sig_ind = apply(MCMC_result_empirical_2$E_omega_trace[[k]][[2]][,z,-1], 2, function(x){prod(hdi(x, 0.7))}) > 0
    significant_omega = rbind(significant_omega, c(k, z, sig_ind))
  }
}
significant_omega = as.data.frame(significant_omega)
colnames(significant_omega) = c("attribute", "forward", "FAST_avg", "FASTWORKS_AVG_ATTEND", "P_RE_HISP", "PHOENIX")
```

```{r}
significant_omega %>%
  group_by(attribute, forward) %>%
  summarise(across(everything(), sum))
```


```{r}
load("C:/Users/Dan Chen/OneDrive/Desktop/BBVI/SCCD cleaned data/TDCM_empirical_sig.RData")
Q_matrix= data$Q_matrix
Y = data$Y
group = data$group
X_group = data$X_group
X_ind = data$X_ind
J = dim(Q_matrix)[1] # the number of questions: 25
K = dim(Q_matrix)[2] # the number of skills: 5
N = dim(Y)[1] # the number of respondents: 987
indT = dim(Y)[2] # the number of time points: 2
G = max(group) # the number of groups: 52
p = dim(X_group[[1]][[1]])[2]  # number of group covariates
m = dim(X_ind[[1]][[2]])[2]  # number of individual covariates (including intercepts)

empirical_data = list('Y' = Y, 
                      'X_group' = X_group, 
                      'X_ind' = X_ind, 
                      'group' = group, 
                      'Q' = Q_matrix, 
                      'K' = K,
                      'indT' = indT
                      )

MCMC_result_empirical = MCMC_Ind_fit(empirical_data, M = 2000, priors=list(beta_prior = 1, sigma_omega = 2.5, v_prior = 1), beta_interact = FALSE, two_way_only = TRUE)


saveRDS(MCMC_result_empirical, file = "MCMC_result_empirical_4.rds")
```
```{r}
MCMC_result_empirical_4 = readRDS("MCMC_result_empirical_4.rds")
```

```{r}
k = 1
z = 1

group_ind_interact = as.data.frame(MCMC_result_empirical_4$E_omega_trace[[k]][[2]][,z,4,-1]) 
colnames(group_ind_interact) = c("FAST_avg", "FASTWORKS_AVG_ATTEND", "P_RE_HISP", "PHOENIX")
group_ind_interact = group_ind_interact %>%
  pivot_longer(cols = everything(), names_to = "Individual variable", values_to = "Coefficient")
   
```


```{r}
ggplot(group_ind_interact, aes(x = `Individual variable`, y = Coefficient)) + 
  geom_violin() +
  theme_bw()
```

```{r}
sig_omega_gamma = matrix(ncol = 7)
for(k in 1:K) {
  for(z in 1:2) {
    for(g in 1:6) {
      sig_ind = apply(MCMC_result_empirical_4$E_omega_trace[[k]][[2]][,z,g,-1], 2, function(x){prod(hdi(x, 0.80))}) >= 0
      sig_omega_gamma = rbind(sig_omega_gamma, c(k, z, g, sig_ind))
    }
  }
}

sig_omega_gamma = as.data.frame(sig_omega_gamma)
colnames(sig_omega_gamma) = c("attribute", "forward", "Ind_var", "FAST_avg", "FASTWORKS_AVG_ATTEND", "P_RE_HISP", "PHOENIX")

sig_omega_gamma %>%
  group_by(attribute, Ind_var) %>%
  summarise(across(everything(), sum))
```

## Interactions identified at 90% credible interval

Hyperactivity (1), Special Education status (5), Hispanic population level (4)
Emotional (2), Special Education status (5), FASTWORKS (3) and Hispanic population level (4)
Conduct (3), Special Education status (5), FASTWORK (3)
Peer (4), English Language Learner (4), Hispanic population level (4)
Prosocial (5), English Language Learner (4), Hispanic population level (4)

```{r}
k = 2
z = 1

school_diffs = matrix(rep(0, 52 * 52), ncol = 52)
for(x in 1:52) {
  for(y in 1:52) {
    school_x = MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,z,x,5]
    school_y = MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,z,y,5]
    school_diffs[x, y] = prod(hdi(school_x - school_y, 0.90)) > 0
  }
}

heatmap(school_diffs, symm = TRUE)
```
```{r}
x = 24
y = 35
school_x = MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,z,x,5]
school_y = MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,z,y,5]
hist(school_x - school_y, breaks = 30)
```

```{r}
plot_interaction = function(model, data, attribute, ind_covariate, main) {
  k = attribute
  g = ind_covariate
  
  data_interaction = cbind(data$X_group[[k]][[2]][,"FASTWORKS_AVG_ATTEND"], colMeans(model$E_gamma_trace[[k]][[2]][,1,,g]))
  data_interaction = as.data.frame(rbind(data_interaction, cbind(data$X_group[[k]][[2]][,"FASTWORKS_AVG_ATTEND"], colMeans(model$E_gamma_trace[[k]][[2]][,2,,g]))))
  colnames(data_interaction) = c("FASTWORK average attendance", "transition coefficient")
  #data_interaction$location = rep(ifelse(data$X_group[[k]][[2]][,"PHOENIX"] == 1, "San Antonio", "Phoenix"), 2)
  data_interaction$`Hispanic population level` = as.factor(data$X_group[[k]][[2]][,"P_RE_HISP"])
  data_interaction$transition = rep(c("01", "11"), each = 52)
  ggplot(data_interaction, aes(x = `FASTWORK average attendance`, y = `transition coefficient`, col = `Hispanic population level`)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~transition) +
    theme_bw() +
    ggtitle(main)
}
```




```{r}
k1_sped = plot_interaction(MCMC_result_empirical_4, data, 1, 5, "Hyperactivity")
k1_sped
```
```{r}
k2_sped = plot_interaction(MCMC_result_empirical_4, data, 2, 5, "Emotional")
k2_sped
```
```{r}
k3_sped = plot_interaction(MCMC_result_empirical_4, data, 3, 5, "Conduct")
k3_sped
```
```{r}
k4_ell = plot_interaction(MCMC_result_empirical_4, data, 4, 4, "Peer")
k4_ell
```
```{r}
k5_ell = plot_interaction(MCMC_result_empirical_4, data, 5, 4, "Prosocial")
k5_ell
```
## Interactions identified at 80% credible interval

Emotional (2), English frequency (2), FASTWORK (3)
Emotional (2), Free or reduced price lunch (3), Hispanic population level (4)
Prosocial (5), Parents attend events (6), FASTWORK (3)

```{r}
plot_interaction(MCMC_result_empirical_4, data, 2, 2, "Mean transition coefficients of english speaking frequency \nfor Emotional attribute")
ggsave("plots/english_emotional.pdf", width = 8, height = 6)
```
```{r}
plot_interaction(MCMC_result_empirical_4, data, 2, 3, "Mean transition coefficients of free or reduced price lunch status \nfor Emotional attribute")
ggsave("plots/reduce_price_emotional.pdf", width = 8, height = 6)
```
```{r}
plot_interaction(MCMC_result_empirical_4, data, 5, 6, "Mean transition coefficients of parental participation in school activities \nfor Prosocial attribute")
ggsave("plots/parent_pprosocial.pdf", width = 8, height = 6)
```

## Direct effects of group level covariate
```{r}
k1_intercept = plot_interaction(MCMC_result_empirical_4, data, 1, 1, "Hyperactivity")
```

```{r}
k2_intercept = plot_interaction(MCMC_result_empirical_4, data, 2, 1, "Emotional")
```

```{r}
k3_intercept = plot_interaction(MCMC_result_empirical_4, data, 3, 1, "Conduct")
```

```{r}
k4_intercept = plot_interaction(MCMC_result_empirical_4, data, 4, 1, "Peer")
```

```{r}
k5_intercept = plot_interaction(MCMC_result_empirical_4, data, 5, 1, "Prosocial")
```




```{r}
k = 4
g = 1

data_FASTWORK_and_LUNCH = cbind(data$X_group[[k]][[2]][,"FAST_avg"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,1,,g]))
data_FASTWORK_and_LUNCH = as.data.frame(rbind(data_FASTWORK_and_LUNCH, cbind(data$X_group[[k]][[2]][,"FAST_avg"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,2,,g]))))
colnames(data_FASTWORK_and_LUNCH) = c("FASTWORK average attendance", "transition coefficient")
data_FASTWORK_and_LUNCH$phoenix = rep(ifelse(data$X_group[[k]][[2]][,"PHOENIX"] == 1, "San Antonio", "Phoenix"), 2)
data_FASTWORK_and_LUNCH$transition = rep(c("forward", "backward"), each = 52)
ggplot(data_FASTWORK_and_LUNCH, aes(x = `FASTWORK average attendance`, y = `transition coefficient`, col = phoenix)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~transition) +
  theme_bw() +
  ggtitle("Mean transition model coefficients of free or reduced price lunch status \nby FASTWORK average attendance for Conduct attribute")
```

```{r}
k = 1
g = 3

data_P_RE_HISP_and_ELL = cbind(data$X_group[[k]][[2]][,"P_RE_HISP"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,1,,g]))
data_P_RE_HISP_and_ELL = as.data.frame(rbind(data_P_RE_HISP_and_ELL, cbind(data$X_group[[k]][[2]][,"P_RE_HISP"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,2,,g]))))
colnames(data_P_RE_HISP_and_ELL) = c("Hispanic population level", "transition coefficient")
data_P_RE_HISP_and_ELL$phoenix = rep(ifelse(data$X_group[[k]][[2]][,"PHOENIX"] == 1, "San Antonio", "Phoenix"), 2)
data_P_RE_HISP_and_ELL$transition = rep(c("forward", "backward"), each = 52)
ggplot(data_P_RE_HISP_and_ELL, aes(x = phoenix, y = `transition coefficient`, col = phoenix)) +
  geom_boxplot(aes(group = phoenix)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~transition + `Hispanic population level`) +
  theme_bw() +
  ggtitle("Mean transition model coefficients of free or reduced price lunch status \nby FASTWORK average attendance for Hyperactivity attribute") +
  xlab(NULL)
```


```{r}
k = 1
g = 4

data_P_RE_HISP_and_ELL = cbind(data$X_group[[k]][[2]][,"P_RE_HISP"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,1,,g]))
data_P_RE_HISP_and_ELL = as.data.frame(rbind(data_P_RE_HISP_and_ELL, cbind(data$X_group[[k]][[2]][,"P_RE_HISP"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,2,,g]))))
colnames(data_P_RE_HISP_and_ELL) = c("Hispanic population level", "transition coefficient")
data_P_RE_HISP_and_ELL$phoenix = rep(ifelse(data$X_group[[k]][[2]][,"PHOENIX"] == 1, "San Antonio", "Phoenix"), 2)
data_P_RE_HISP_and_ELL$transition = rep(c("forward", "backward"), each = 52)
ggplot(data_P_RE_HISP_and_ELL, aes(x = phoenix, y = `transition coefficient`, col = phoenix)) +
  geom_boxplot(aes(group = phoenix)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~transition + `Hispanic population level`) +
  theme_bw() +
  ggtitle("Mean transition model coefficients of free or reduced price lunch status \nby FASTWORK average attendance for Hyperactivity attribute")
```


```{r}
k = 1

data_P_RE_HISP_and_ELL = cbind(as.factor(data$X_group[[k]][[2]][,"P_RE_HISP"]), colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,1,,4]))
data_P_RE_HISP_and_ELL = rbind(data_P_RE_HISP_and_ELL, cbind(as.factor(data$X_group[[k]][[2]][,"P_RE_HISP"]), colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,2,,4])))
data_P_RE_HISP_and_ELL = as.data.frame(data_P_RE_HISP_and_ELL)
colnames(data_P_RE_HISP_and_ELL) = c("Hispanic population level", "transition coefficient")
data_P_RE_HISP_and_ELL$transition = rep(c("forward", "backward"), each = 52)
data_P_RE_HISP_and_ELL$phoenix = rep(ifelse(data$X_group[[k]][[2]][,"P_RE_HISP"] == 1, "San Antonio", "Phoenix"), 2)

ggplot(data_P_RE_HISP_and_ELL, aes(x = `Hispanic population level`, y = `transition coefficient`)) +
  geom_boxplot(aes(group = `Hispanic population level`)) +
  geom_point() +
  facet_wrap(~transition) +
  theme_bw() +
  ggtitle("Mean transition model coefficients of english language learner status \ngrouped by level of hispanic population for Hyperactivity attribute")
```

```{r}
k = 3

data_FASTWORK_and_SPED = cbind(data$X_group[[k]][[2]][,"FASTWORKS_AVG_ATTEND"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,1,,5]))
data_FASTWORK_and_SPED = as.data.frame(rbind(data_FASTWORK_and_SPED, cbind(data$X_group[[k]][[2]][,"FASTWORKS_AVG_ATTEND"], colMeans(MCMC_result_empirical_4$E_gamma_trace[[k]][[2]][,2,,5]))))
colnames(data_FASTWORK_and_SPED) = c("FASTWORK average attendance", "transition coefficient")
data_FASTWORK_and_SPED$phoenix = rep(ifelse(data$X_group[[k]][[2]][,"PHOENIX"] == 1, "San Antonio", "Phoenix"), 2)
data_FASTWORK_and_SPED$transition = rep(c("forward", "backward"), each = 52)
ggplot(data_FASTWORK_and_SPED, aes(x = `FASTWORK average attendance`, y = `transition coefficient`, col = phoenix)) +
  geom_point() +
  facet_wrap(~transition) +
  theme_bw() +
  ggtitle("Mean transition model coefficients of special eduation status \nby FASTWORK average attendance for Conduct attribute")
```

```{r}
sped_interact = k1_sped + xlab(NULL) + theme(legend.position = "none") + ylim(-2.7, 2.1) +
  k2_sped + xlab(NULL) + ylab(NULL) + theme(legend.position = "none") + ylim(-2.7, 2.1) +
  k3_sped + xlab(NULL) + ylab(NULL) + ylim(-2.7, 2.1) +
  plot_annotation(title = "Mean transition coefficient for special education status") &
  xlab("FASTWORKS average attendance")
ggsave("plots/SPED interactions.pdf", plot = sped_interact, width = 12, height = 4)
```


```{r}
ell_interact = k4_ell + ylim(-1.1, 1) + theme(legend.position = "none") +
  k5_ell + ylab(NULL) + ylim(-1.1, 1) +
  plot_annotation("Mean transition coefficient for english language learner status")
ggsave("plots/ELL interactions.pdf", plot = ell_interact, width = 10, height = 4)
```

```{r}
intercepts = (k1_intercept + theme(legend.position = "none") |
  k2_intercept + theme(legend.position = "none") + ylab(NULL) |
  k3_intercept + theme(legend.position = "none") + ylab(NULL)) /
  (k4_intercept + theme(legend.position = "none") |
  k5_intercept + ylab(NULL) |
  plot_spacer()) +
  plot_annotation(title = "Mean transition intercepts")
ggsave("plots/intercepts.pdf", plot = intercepts, height = 6, width = 10)
```

```{r}
intercepts
```




```{r}
s4t2_f_SEPD = as.data.frame(MCMC_result_empirical_4$E_gamma_trace[[4]][[2]][,2,,6]) %>%
  pivot_longer(cols = everything(), names_to = "School", values_to = "Coefficient")
s4t2_f_SEPD$School = as.factor(substr(s4t2_f_SEPD$School, 2, 4))

ggplot(s4t2_f_SEPD, aes(x = School, y = Coefficient)) +
  geom_violin(scale = "width", color = "transparent", fill = "seagreen") +
  theme_bw() +
  labs(title = "Effect of special education status on acquiring attribute 2") +
  scale_x_discrete(labels = NULL)
```


```{r}
k1_parent = plot_interaction(MCMC_result_empirical_4, data, 1, 6, "Hyperactivity")
k2_parent = plot_interaction(MCMC_result_empirical_4, data, 2, 6, "Emotional")
k3_parent = plot_interaction(MCMC_result_empirical_4, data, 3, 6, "Conduct")
k4_parent = plot_interaction(MCMC_result_empirical_4, data, 4, 6, "Peer")
k5_parent = plot_interaction(MCMC_result_empirical_4, data, 5, 6, "Prosocial")

parent = (k1_parent + theme(legend.position = "none") + ylim(-0.4, 0.5) |
  k2_parent + theme(legend.position = "none") + ylab(NULL) + ylim(-0.4, 0.5) |
  k3_parent + theme(legend.position = "none") + ylab(NULL) + ylim(-0.4, 0.5) |
  k4_parent + theme(legend.position = "none") + ylab(NULL) + ylim(-0.4, 0.5) |
  k5_parent + ylab(NULL) + ylim(-0.4, 0.5)) + plot_annotation(title = "Mean transition coefficients for parents attending school activities")
parent
```

```{r}
k1_english = plot_interaction(MCMC_result_empirical_4, data, 1, 2, "Hyperactivity")
k2_english = plot_interaction(MCMC_result_empirical_4, data, 2, 2, "Emotional")
k3_english = plot_interaction(MCMC_result_empirical_4, data, 3, 2, "Conduct")
k4_english = plot_interaction(MCMC_result_empirical_4, data, 4, 2, "Peer")
k5_english = plot_interaction(MCMC_result_empirical_4, data, 5, 2, "Prosocial")

english = (k1_english + theme(legend.position = "none") + ylim(-0.75, 0.5) |
  k2_english + theme(legend.position = "none") + ylab(NULL) + ylim(-0.75, 0.5) |
  k3_english + theme(legend.position = "none") + ylab(NULL) + ylim(-0.75, 0.5) |
  k4_english + theme(legend.position = "none") + ylab(NULL) + ylim(-0.75, 0.5) |
  k5_english + ylab(NULL) + ylim(-0.75, 0.5)) + plot_annotation(title = "Mean transition coefficients for english speaking frequency")
english
```

```{r}
k1_lunch = plot_interaction(MCMC_result_empirical_4, data, 1, 3, "Hyperactivity")
k2_lunch = plot_interaction(MCMC_result_empirical_4, data, 2, 3, "Emotional")
k3_lunch = plot_interaction(MCMC_result_empirical_4, data, 3, 3, "Conduct")
k4_lunch = plot_interaction(MCMC_result_empirical_4, data, 4, 3, "Peer")
k5_lunch = plot_interaction(MCMC_result_empirical_4, data, 5, 3, "Prosocial")

lunch = (k1_lunch + theme(legend.position = "none") + ylim(-0.6, 0.6) |
  k2_lunch + theme(legend.position = "none") + ylab(NULL) + ylim(-0.6, 0.6) |
  k3_lunch + theme(legend.position = "none") + ylab(NULL) + ylim(-0.6, 0.6) |
  k4_lunch + theme(legend.position = "none") + ylab(NULL) + ylim(-0.6, 0.6) |
  k5_lunch + ylab(NULL) + ylim(-0.6, 0.6)) + plot_annotation(title = "Mean transition coefficients for free or reduced price lunch status")
lunch
```
```{r}
k1_ell = plot_interaction(MCMC_result_empirical_4, data, 1, 4, "Hyperactivity")
k2_ell = plot_interaction(MCMC_result_empirical_4, data, 2, 4, "Emotional")
k3_ell = plot_interaction(MCMC_result_empirical_4, data, 3, 4, "Conduct")
k4_ell = plot_interaction(MCMC_result_empirical_4, data, 4, 4, "Peer")
k5_ell = plot_interaction(MCMC_result_empirical_4, data, 5, 4, "Prosocial")

ell = (k1_ell + theme(legend.position = "none") + ylim(-1.5, 1) |
  k2_ell + theme(legend.position = "none") + ylab(NULL) + ylim(-1.5, 1) |
  k3_ell + theme(legend.position = "none") + ylab(NULL) + ylim(-1.5, 1) |
  k4_ell + theme(legend.position = "none") + ylab(NULL) + ylim(-1.5, 1) |
  k5_ell + ylab(NULL) + ylim(-1.5, 1)) + plot_annotation(title = "Mean transition coefficients for english language learner status")
ell
```
```{r}
k1_sped = plot_interaction(MCMC_result_empirical_4, data, 1, 5, "Hyperactivity")
k2_sped = plot_interaction(MCMC_result_empirical_4, data, 2, 5, "Emotional")
k3_sped = plot_interaction(MCMC_result_empirical_4, data, 3, 5, "Conduct")
k4_sped = plot_interaction(MCMC_result_empirical_4, data, 4, 5, "Peer")
k5_sped = plot_interaction(MCMC_result_empirical_4, data, 5, 5, "Prosocial")

sped = (k1_sped + theme(legend.position = "none") + ylim(-2.75, 2.1) |
  k2_sped + theme(legend.position = "none") + ylab(NULL) + ylim(-2.75, 2.1) |
  k3_sped + theme(legend.position = "none") + ylab(NULL) + ylim(-2.75, 2.1) |
  k4_sped + theme(legend.position = "none") + ylab(NULL) + ylim(-2.75, 2.1) |
  k5_sped + ylab(NULL) + ylim(-2.75, 2.1)) + plot_annotation(title = "Mean transition coefficients for special education status")
sped
```

```{r}
intercepts = (k1_intercept + theme(legend.position = "none")  + ylim(-1.7, 4.1) |
  k2_intercept + theme(legend.position = "none") + ylab(NULL)  + ylim(-1.7, 4.1) |
  k3_intercept + theme(legend.position = "none") + ylab(NULL)  + ylim(-1.7, 4.1)) /
  (k4_intercept + theme(legend.position = "none")  + ylim(-1.7, 4.1) |
  k5_intercept + ylab(NULL)  + ylim(-1.7, 4.1) |
  plot_spacer()) +
  plot_annotation(title = "Transition intercepts posterior means")
ggsave("plots/intercepts.pdf", plot = intercepts, height = 6, width = 10)
```

```{r}
intercepts
```







